---
title: "Classification"
subtitle: "Part 3"
author: "Prof. Bisbee"
institute: "Vanderbilt University"
date: "Lecture Date: 2022/11/28\n Slides Updated: `r Sys.Date()`"
output:
  xaringan::moon_reader:
    # self_contained: true
    chakra: libs/remark-latest.min.js
    lib_dir: libs
    css:
      - default
      - css/lexis.css
      - css/lexis-fonts.css
    #seal: false
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
      #ratio: "16:9"

---

```{css,echo = F}
.small .remark-code { /*Change made here*/
  font-size: 85% !important;
}
.tiny .remark-code { /*Change made here*/
  font-size: 50% !important;
}
```

```{r,include=F}
options(width=60)
knitr::opts_chunk$set(fig.align='center',fig.width=9,fig.height=5)
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})
```

```{r klippy, echo=FALSE, include=TRUE}
klippy::klippy()
```

# Agenda

1. Recap of classification

2. Recap of evaluation

2. Using classifiers

---

# Classification Recap

- Similar to **prediction**

--

- Difference is in the **outcome variable** $Y$:

--

  - Continuous: use regression model
  
  - Categorical / Binary: use classification model
  
--

- **NB:** *classification* is a type of *prediction*

---

# Classification Recap

- Thus far, only used binary $Y$

--

  - *Can* use linear regression model
  
  - *Better* to use logistic regression model
  
--

- Linear regression: `lm(formula,data)`

--

  - Predicts "latent" measures (`predict(mLM)`)
  
  - I.e., willingness to attend
  
--

- Logistic: `glm(formula,data,family = binomial(link = 'logit'))`

--

  - Predicts probabilities (`predict(mLG,type = 'response')`)
  
  - I.e., probability of attending
  
```{r,message=F}
require(tidyverse)
require(tidymodels)
ad <- readRDS('../data/admit_data.rds')
```

---

# Classification Recap

```{r,message = F}
ad %>% ggplot(aes(x = sat,y = yield)) + 
  geom_jitter(width = .01,height = .05,alpha = .25) + 
  geom_smooth(method = 'lm',color = 'black') + 
  geom_smooth(method = 'glm',color = 'red',
              method.args = list(family = binomial(link = 'logit')))
```

---

# Classification Recap

```{r,message = F,warning = F}
ad %>% ggplot(aes(x = income,y = yield)) + 
  geom_jitter(width = .01,height = .05,alpha = .25) + 
  geom_smooth(method = 'lm',color = 'black') + 
  geom_smooth(method = 'glm',color = 'red',
              method.args = list(family = binomial(link = 'logit')))
```

---

# Classification Recap

```{r,message = F,warning = F}
ad %>% ggplot(aes(x = log(income),y = yield)) + 
  geom_jitter(width = .01,height = .05,alpha = .25) + 
  geom_smooth(method = 'lm',color = 'black') + 
  geom_smooth(method = 'glm',color = 'red',
              method.args = list(family = binomial(link = 'logit')))
```


---

# Classification Workflow

1. Train: `mLG <- glm(formula,data,family = binomial(link = 'logit'))`

2. Predict: `data$predY <- predict(mLG,type = 'response')`

3. Evaluate: `roc_auc(data,truth,estimate)`

--

- **NB:** Evaluation stage is *DEEP*

--

  - Classify observations based on threshold: `ifelse(predY > 0.5,1,0)`
  
  - Calculate accuracy by group across thresholds
  
  - **Sensitivity** and **Specificity** &rarr; ROC curve &rarr; AUC
  
  - Should be estimated via cross validation to prevent **overfitting**
  
  - Make sure `truth` is ordered in **reverse**: `factor(truth,levels = c('1','0'))`


---

# Classification Workflow

- What do we get out of all this effort?

--

- A single classification algorithm that we think is best

--

.small[
```{r,message=F,warning=F}
# Train
form <- 'yield ~ sat + legacy + visit + registered + sent_scores + income + gpa + distance + net_price'
mLG <- glm(formula = as.formula(form),
           data = ad,family = binomial(link = 'logit'))
# Predict
pred <- ad %>%
  mutate(predY = predict(mLG,type = 'response'), # NB: type = 'response' for glm()!
         truth = factor(yield,levels = c('1','0'))) # NB: reorder outcome so that 1 is first!
# Evaluate
roc_auc(data = pred,truth = 'truth',estimate = 'predY')
```
]


---

# Classification Workflow

- But we should actually do this with **cross validation**

--

.small[
```{r,message = F,warning = F}
cvRes <- NULL
for(i in 1:100) {
  inds <- sample(1:nrow(ad),size = round(nrow(ad)*.8),replace = F)
  train <- ad %>% slice(inds)
  test <- ad %>% slice(-inds)
  
  # Train
  mLM <- lm(form,data = train)
  mLG <- glm(form,data = train,family = binomial(link = 'logit'))
  # Predict
  pred <- test %>%
  mutate(predLM = predict(mLM,newdata = test),
         predLG = predict(mLG,newdata = test,type = 'response'),
         truth = factor(yield,levels = c('1','0')))
  # Evaluate
  resLG <- roc_auc(data = pred,truth = 'truth',estimate = 'predLG') %>%
    mutate(algo = 'logit')
  resLM <- roc_auc(data = pred,truth = 'truth',estimate = 'predLM') %>%
    mutate(algo = 'linear') 
  cvRes <- resLG %>% bind_rows(resLM) %>% bind_rows(cvRes)
}
```
]

---

# Classification Workflow

```{r}
cvRes %>%
  ggplot(aes(x = .estimate,fill = algo)) + 
  geom_density(alpha = .4)
```

---

# Using the Algorithm

- The algorithm can help us make decisions

--

- Recall the goals and constraints of this college:

--

  - **Goals**: Increase SAT to 1300
  
  - **Goals**: Admit 200 more students with incomes under $50,000
  
  - **Constraints**: Maintain total revenues of at least $30m
  
  - **Constraints**: Maintain entering class size of at least 1,466
  
--

- Tools: 

--

  - Need-based and Merit-based aid
  
--

  - Effort for visit and registration
  
--

  - Other targeting? (ethics?)

---

# Using the Algorithm

1. Hypotheticals for specific types of students

--

  - What is the probability a given student will attend?
  
--

2. Hypotheticals for changes in policy

--

  - If we increase price, what will happen to attendance?
  
--

- `data_grid()` function from `modelr` package can help!
  
---

# Using the Algorithm

`1.` What is the probability that this student will attend?
```{r,message = F,warning = F}
mLGFinal <- glm(formula = as.formula(form),
           data = ad,family = binomial(link = 'logit'))
require(modelr)
hypo_data <- ad %>%
  data_grid(legacy = 0,
            visit = 1,
            registered = 1,
            sent_scores = 1,
            sat = 1400,
            income = 95000,
            gpa = 3.9,
            distance = .1,
            net_price = 6875)

predict(mLGFinal,newdata = hypo_data,type = 'response')
```

---

# Using the Algorithm

`2.` What is the probability they will attend if we increase the price by $10k?
```{r,message = F}
hypo_data <- ad %>%
  data_grid(legacy = 0,
            visit = 1,
            registered = 1,
            sent_scores = 1,
            sat = 1400,
            income = 95000,
            gpa = 3.9,
            distance = .1,
            net_price = 16875)

predict(mLGFinal,newdata = hypo_data,type = 'response')
```

---

# Using the Algorithm

- Can combine with `data_grid()`

```{r,message = F}
hypo_data <- ad %>%
  data_grid(legacy = 0,
            visit = 1,
            registered = 1,
            sent_scores = 1,
            sat = 1400,
            income = 95000,
            gpa = 3.9,
            distance = .1,
            net_price = c(6875,16875))

predict(mLGFinal,newdata = hypo_data,type = 'response')
```

---

# Using the Algorithm

- Can use `data_grid()` to calculate "typical" values for every variable

--

  - **Typical**: Mean for continuous, mode for categorical
  
--

```{r}
(hypo_data <- ad %>%
  data_grid(.model = mLGFinal))

predict(mLGFinal,newdata = hypo_data,type = 'response')
```

---

# Using the Algorithm

- Compare otherwise typical admits who differ only in terms of GPA

--

```{r}
hypo_data <- ad %>%
  data_grid(.model = mLGFinal,
            gpa = c(3.5,3.9))

predict(mLGFinal,newdata = hypo_data,type = 'response')
```

---

# Using the Algorithm

- Multiple comparisons are possible

--

```{r,message = F}
require(ggridges)
hypo_data <- ad %>%
  data_grid(.model = mLGFinal,
            gpa = seq_range(gpa,n = 100),
            legacy = c(0,1))

toplot <- hypo_data %>%
  mutate(preds = predict(mLGFinal,newdata = hypo_data,type = 'response'))

p <- toplot %>%
  ggplot(aes(x = gpa,y = preds,color = factor(legacy))) + 
  geom_line() + 
  theme_ridges() + 
  scale_y_continuous(labels = scales::percent) + 
  labs(title = 'Predicted Attendance by GPA and Legacy',
       x = 'GPA',
       y = 'Prob(Attend)')
```

---

# Using the Algorithm

```{r}
p
```

---

# Consulting Tasks

- Currently, admit 1466 at ~$30.7m revenues

```{r}
ad %>%
  filter(yield == 1) %>%
  summarise(totRev = dollar(sum(net_price)),
            totAtt = n())
```

---

# Consulting Tasks

- What if we increased the price for those who submit scores?

--

```{r}
# Currently, 355 students who sent scores are predicted to attend
ad %>%
  mutate(preds = predict(mLGFinal,type = 'response')) %>%
  mutate(pred_class = ifelse(preds > .5,1,0)) %>%
  count(sent_scores,pred_class)
```

---

# Consulting Tasks

- What if we increased the price for those who submit scores?

```{r}
# If price increases by $5k, number of predicted drops to 338
hypo <- ad %>%
  mutate(net_price = ifelse(sent_scores == 1,
                            net_price + 5000,net_price))

hypo %>%
  mutate(preds = predict(mLGFinal,newdata = hypo,type = 'response')) %>%
  mutate(pred_class = ifelse(preds > .5,1,0)) %>%
  count(sent_scores,pred_class)
```

---

# Consulting Tasks

- What if we increased the price for those who submit scores?

```{r}
# BUT we make more total revenue
hypo %>%
  mutate(preds = predict(mLGFinal,newdata = hypo,type = 'response')) %>%
  mutate(pred_class = ifelse(preds > .5,1,0)) %>%
  filter(pred_class == 1) %>%
  summarise(tot_rev = scales::dollar(sum(net_price)))
```

---

# Consulting Tasks

- What if we increased the price for those who submit scores?

```{r}
# Although total admits declined
hypo %>%
  mutate(preds = predict(mLGFinal,newdata = hypo,type = 'response')) %>%
  mutate(pred_class = ifelse(preds > .5,1,0)) %>%
  filter(pred_class == 1) %>%
  count()
```

---

# Consulting Tasks

- Can we increase SAT scores to 1300?

--

```{r}
ad %>%
  filter(yield == 1) %>%
  summarise(satAvg = round(mean(sat)))
```

--

- Reduce price for those above 1300

---

# Consulting Tasks

- Can we increase SAT scores to 1300?

- Reduce price for those above 1300

```{r}
hypo <- ad %>%
  mutate(net_price = ifelse(sat > 1300,
                            net_price - 5000,net_price))

hypo %>%
  mutate(preds = predict(mLGFinal,newdata = hypo,type = 'response')) %>%
  mutate(pred_class = ifelse(preds > .5,1,0)) %>%
  filter(pred_class == 1) %>%
  summarise(satAvg = round(mean(sat)),
            tot_rev = scales::dollar(sum(net_price)),
            totAttend = n())
```

---

# Consulting Tasks

- Can we increase SAT scores to 1300?

- BUT need to make up losses

```{r}
hypo <- ad %>%
  mutate(net_price = ifelse(sat > 1300,net_price - 5000,
                            ifelse(sat < 1300,net_price + 2500,net_price)))

hypo %>%
  mutate(preds = predict(mLGFinal,newdata = hypo,type = 'response')) %>%
  mutate(pred_class = ifelse(preds > .5,1,0)) %>%
  filter(pred_class == 1) %>%
  summarise(satAvg = round(mean(sat)),
            tot_rev = scales::dollar(sum(net_price)),
            totAttend = n())
```

---

# Consulting Tasks

- Other tools of influence?

```{r}
hypo <- ad %>%
  mutate(net_price = ifelse(sat > 1300,
                            net_price - 2500,
                            ifelse(sat < 1300,
                                   net_price + 5000,net_price)),
         visit = ifelse(sat > 1300,
                            1,
                            ifelse(sat < 1300,
                                   0,visit)),
         registered = ifelse(sat > 1300,
                            1,
                            ifelse(sat < 1300,
                                   0,registered)))
```

---

# Consulting Tasks

- Other tools of influence?

```{r}
hypo %>%
  mutate(preds = predict(mLGFinal,newdata = hypo,type = 'response')) %>%
  mutate(pred_class = ifelse(preds > .5,1,0)) %>%
  filter(pred_class == 1) %>%
  summarise(satAvg = round(mean(sat)),
            tot_rev = scales::dollar(sum(net_price)),
            totAttend = n())
```



---

# Consulting Tasks

- Even with an algorithm, this is hard!

--

- Can use the algorithm to target more precisely

--

.small[
```{r}
hypo <- ad %>%
  mutate(preds = predict(mLGFinal,type = 'response')) %>%
  mutate(ndg = ifelse((preds < .5 & preds > .3),'Nudge Up',
                         ifelse((preds > .5 & preds < .7),'Nudge Down','Ignore'))) %>%
  mutate(net_price = ifelse(sat > 1300 & ndg == 'Nudge Up',
                            net_price - 15000,
                            ifelse(sat < 1300 & ndg == 'Nudge Down',
                                   net_price + 2500,net_price)))

hypo %>%
  mutate(preds = predict(mLGFinal,newdata = hypo,type = 'response')) %>%
  mutate(pred_class = ifelse(preds > .5,1,0)) %>%
  filter(pred_class == 1) %>%
  summarise(satAvg = round(mean(sat)),
            tot_rev = scales::dollar(sum(net_price)),
            totAttend = n())
```
]

---

# Consulting Tasks

- Can use the algorithm to target more precisely

- Target those who are just below the threshold

```{r}
hypo %>%
  mutate(preds = predict(mLGFinal,newdata = hypo,type = 'response')) %>%
  mutate(pred_class = ifelse(preds > .5,1,0)) %>%
  filter(pred_class == 1) %>%
  summarise(satAvg = round(mean(sat)),
            tot_rev = scales::dollar(sum(net_price)),
            totAttend = n())
```

---

# Consulting Tasks

- Or we could target on other variables

--

  - But remember the concerns of **ethics**!
  
--

.small[
```{r}
# summary(mLGFinal)
hypo <- ad %>%
 mutate(net_price = ifelse(sat > 1300,
                            net_price - 10000,
                            ifelse(sat < 1300,
                                   net_price + 10000,net_price)),
         visit = ifelse(sat > 1300,
                            1,
                            ifelse(sat < 1300,
                                   0,visit)),
         registered = ifelse(sat > 1300,
                            1,
                            ifelse(sat < 1300,
                                   0,registered)),
        legacy = ifelse(sat > 1300,1,legacy),
        gpa = ifelse(sat > 1300,3.9,gpa),
        income = ifelse(sat > 1300,150000,income),
        distance = ifelse(sat > 1300,100,distance))
```
]
---

# Consulting Tasks

- Or we could target on other variables

  - But remember the concerns of **ethics**!
  
```{r}
hypo %>%
  mutate(preds = predict(mLGFinal,newdata = hypo,type = 'response')) %>%
  mutate(pred_class = ifelse(preds > .5,1,0)) %>%
  filter(pred_class == 1) %>%
  summarise(satAvg = round(mean(sat)),
            tot_rev = scales::dollar(sum(net_price)),
            totAttend = n())
```

---

# Is this even possible?

```{r}
ad %>%
  mutate(pred_class = ifelse(sat < 1231,0,1)) %>% # Don't admit anyone with an SAT score less than 1231.
  filter(pred_class == 1) %>%
  summarise(satAvg = round(mean(sat)),
            tot_rev = scales::dollar(sum(net_price)),
            totAttend = n())
```

---

# Conclusion

- Homework: can you achieve the other goal?

--

  - Increase number of low income admits by 200
  
  - Subject to the same constraints of revenue and total admits
  
--

- Remember the workflow:

--

  1. Train: `glm()`
  
  2. Predict: `predict()`
  
  3. Evaluate: `roc_auc()`
  
--

  4. Adjust: `net_price = ifelse()`
